Portable Multi- and Many-Core Performance for Finite-Difference/Element Codes; Application to the Free-Surface Component of NEMO (NEMOLite2D 1.0).

### 2017-09-15 docs/papers/nemolite2d_performance/referee_response.txt

(Top of page 8.) We have clarified why allocating larger arrays was
necessary in order to use the safe_address directive.

Changed "Intel does..." to "Intel can..." as requested.

Sec 2.1.5 - the grid is not attached to the field, rather each field
keeps a pointer back to the grid upon which it is defined. This approach
was chosen in order to support applications which use more than one grid.

Sec 2.1.8 - it is not clear why in-lining manually helps (especially since
one would expect the Intel compiler in particular to do this). We suspect
that the change in the PSy-layer code causes the compiler to make some
different optimisation decisions and these happen to be beneficial. A
proper investigation of this is outside the scope of this paper. We have
added a sentence about this.

Sec 2.2.1 - yes, other strategies might be beneficial, especially blocking.
The key point is that adopting the PSyKAl structure does not limit these
possibilities.

Figure 3 - The Gnu results are much slower because gfortran is only SIMD
vectorising the loop containing the Continuity kernel. For some reason
it does not vectorise any others. Unlike Cray, there is no directive
that can be inserted to mandate that a loop be SIMD vectorised. We have
added text to this effect.

### 2017-09-15 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: respond to more of Carlos comments

The Time-update section includes several array copies where fields for
the current time-step become the fields at the previous time-step.
Initially we implemented these copies as 'built-in' kernels (in the
GOcean infrastructure) as they are specified in the algorithm
layer. However, we obtained better performance by simply in-lining
these array copies into the PSy layer.

@@

Moving now the to the {PS}y{KA}l version of NEMOLite2D,
Figure~\ref{FIG_best_psykal_perf_summary} plots the performance of the
fastest {PS}y{KA}l version for each of the compiler/problem-size combinations.
While the Intel compiler still produces the best-performing binary,
the Cray-compiled binary is now a very close second. In fact, the
performance of both the Gnu- and Cray-compiled {PS}y{KA}l versions is
generally significantly greater than that of their respective original
versions. We also note that best absolute performance (in terms of
grid points processed per second) with any compiler is obtained with
the $256^2$ domain.

@@

Figure~\ref{FIG_slowdown_summary} plots the percentage difference
between the performance of the original and the best {PS}y{KA}l versions
of NEMOLite2D for each compiler/problem-size combination. This shows that it is
only the Intel-compiled binary running the $64^2$ domain that is
slower with the {PS}y{KA}l version of the code (and then only by some
3\%). For all other points in the space, the optimised {PS}y{KA}l version
of the code performs better. The results for the Cray compiler are
however somewhat sqewed by the fact that it did not SIMD vectorise key
parts of the original version (see below).

>>>>

The Time-update section includes several array copies where fields for
the current time-step become the fields at the previous time-step.
Initially we implemented these copies as `built-in' kernels (in the
GOcean infrastructure) as they are specified in the Algorithm
layer. However, we obtained better performance (for the Gnu and Intel
compilers) by simply manually in-lining these array copies into the
PSy layer. As discussed in~\ref{sec_inline_mom}, it is unclear why
in-lining should improve performance, particularly for the Intel
compiler. However, a detailed investigation of this issue is outside
the scope of this paper.

@@

Moving now the to the {PS}y{KA}l version of NEMOLite2D,
Figure~\ref{FIG_best_psykal_perf_summary} plots the performance of the
fastest {PS}y{KA}l version for each of the compiler/problem-size
combinations.  While the Intel compiler still produces the
best-performing binary, the Cray-compiled binary is now a very close
second. In fact, the performance of both the Gnu- and Cray-compiled
{PS}y{KA}l versions is generally significantly greater than that of
their respective original versions. We also note that best absolute
performance (in terms of grid points processed per second) with any
compiler is obtained with the $256^2$ domain. The performance of the
Gnu-compiled binary is consistently a factor of two slower than those
of the other compilers. This is due to the fact that it only SIMD
vectorises the loop performing the Continuity calculation and, unlike
Cray and Intel, it does not have an intrinsic to make vectorisation of
a loop mandatory. (When using OpenMP 4.0 the SIMD directive could be
used but even then that is only taken as a hint.)

@@

Figure~\ref{FIG_slowdown_summary} plots the percentage difference
between the performance of the original and the best {PS}y{KA}l versions
of NEMOLite2D for each compiler/problem-size combination. This shows that it is
only the Intel-compiled binary running the $64^2$ domain that is
slower with the {PS}y{KA}l version of the code (and then only by some
3\%). For all other points in the space, the optimised {PS}y{KA}l version
of the code performs better. The results for the Cray compiler are
however somewhat skewed by the fact that it did not SIMD vectorise key
parts of the original version (see below).
